---
title: "Identifying forged banknote from genuine banknotes using Logistic Regression"
author: "Banknote Authentication"
date: "11/23/2021"
output:
  pdf_document: default
  html_document: default
---

#### Problem 1: Logistic Regression 

This question should be answered using the ***"Banknote Authentication"*** data set. Description about the data set can be found on the link provided. Objective of this question is to fit an logistic regression model to classify forged banknote from genuine banknotes. (Presumably 0 for genuine and 1 for forged bank notes)

a. Produce some numerical and graphical summaries of the data set. Explain the relationships.

```{r}
library(ISLR)
library(GGally)
library(tidyverse)
library("reshape2")
df=read.csv("~/Desktop/banknote_authentication.txt")
names(df)
dim(df)
str(df)

summary (df)

count = df %>%                              
  group_by(class) %>%
  summarise(count = n_distinct(Variance))
count  
```

b. Is this a balanced data set?.
No, it is an imbalanced data

c. Use the full data set to perform a logistic regression with _Class_ as the response variable. Do any of the predictors appear to be statistically significant? If so, which ones?
```{r}
log <- glm(class ~ Variance + skewness + curtosis + entropy, data = df, family = "binomial")
summary(log)

pred=predict(log,type="response")


glm.pred=rep("0",nrow(df))
glm.pred[pred>.5]="1"


table(glm.pred,df$class)

mean(glm.pred==df$class)
```

d. Compute the confusion matrix and overall fraction of correct predictions. Explain what the confusion matrix id telling you about the types of mistakes made by logistic regression.
```{r}
training=sample(c(TRUE,FALSE),size=nrow(df),prob=c(0.8,0.2),replace=TRUE)
df.test=df[!training,]
dim(df.test)
class.test=df$class[!training] 

```

e. Create a training set with 80% of the observations, and a testing set containing the remaining 20%.Compute the confusion matrix and the overall fraction of correct prediction for the testing data set.
```{r}
logReg2=glm(class~.,data=df,family=binomial,subset=training)

test.predicted.m2 <- predict(logReg2, df.test, type = "response")

glm.pred2=rep("0",nrow(df.test))

glm.pred2[test.predicted.m2>.5]="1"

tbl2 = round(table(class.test, glm.pred2)/nrow(df.test),3)
tbl2

m2.error = mean(class.test == glm.pred2)
m2.error

```
