---
title: "Exploratory Data Analysis Project"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE,
                      comment = "")
```

# Introduction

This semester we will be working with a dataset of all domestic outbound flights from Dulles International Airport in 2016.

Airports depend on accurate flight departure and arrival estimates to maintain operations, profitability, customer satisfaction, and compliance with state and federal laws. Flight performance, including departure and arrival delays must be monitored, submitted to the Federal Aviation Agency (FAA) on a regular basis, and minimized to maintain airport operations. **The FAA considered a flight to be delayed if it has an arrival delay of at least 15 minutes**.

The executives at Dulles International Airport have hired you as a Data Science consultant to perform an exploratory data analysis on all domestic flights from 2016 and produce an executive summary of your key insights and recommendations to the executive team.

Before you begin, take a moment to read through the following airline flight terminology to familiarize yourself with the industry: [Airline Flight Terms](https://www.myairlineflight.com)



## Dulles Flights Data

The `flights_df` data frame is loaded below and consists of 33,433 flights from IAD (Dulles International) in 2016. The rows in this data frame represent a single flight with all of the associated features that are displayed in the table below.

**Note**: If you have not installed the `tidyverse` package, please do so by going to the `Packages` tab in the lower right section of RStudio, select the `Install` button and type `tidyverse` into the prompt. If you cannot load the data, then try downloading the latest version of R (at least 4.0). The `readRDS()` function has different behavior in older versions of `R` and may cause loading issues.



```{r}

library(tidyverse)
library(ggplot2)
library(dplyr)

flights_df <- readRDS(url('https://gmubusinessanalytics.netlify.app/data/dulles_flights.rds'))

```


### Raw Data

```{r}

flights_df

delayedFlights_df <- flights_df %>% filter(arrival_delay >= 15)
delayedFlights_df

```


# Exploratory Data Analysis


Executives at this company have hired you as a data science consultant to evaluate their flight data and make recommendations on flight operations and strategies for minimizing flight delays.

You must think of **at least 8 relevant questions** that will provide evidence for your recommendations.

The goal of your analysis should be discovering which variables drive the differences between flights that are early/on-time vs. flights that are delayed.

Some of the many questions you can explore include:

-   Are flight delays affected by taxi-out time? Do certain airlines or time of 
    year lead to greater taxi out times (i.e. traffic jams on the runways)?

-   Are certain times of the day or year problematic?

-   Are certain destination or airlines prone to delays?


You must answer each question and provide supporting data summaries with either a summary data frame (using `dplyr`/`tidyr`) or a plot (using `ggplot`) or both.

In total, you must have a minimum of 5 plots and 4 summary data frames for the exploratory data analysis section. Among the plots you produce, you must have at least 4 different types (ex. box plot, bar chart,
histogram, heat map, etc...)

Each question must be answered with supporting evidence from your tables and plots.


## Question 1



**Question**:Are flight delays affected by taxi-out time?



**Answer**: No, flight delay doesn't seem related/ affected by taxi-out time because if looking from left to right, the taxi out time and airlines time is more around 250 on x-axis and 50 on y-axis which is not moving further. 



To add additional R code chunks for your work, select `Insert` then `R` from the top of this notebook file.





```{r}
geom_smooth(span = 0.15)
ggplot(data = delayedFlights_df, mapping = aes(x =arrival_delay , y = taxi_out)) +
       geom_point() + geom_smooth(span = 15) +  labs(title = "Boxplot of Taxi-out time by Airlines",
       x = "Airlines", 
       y = "Taxi-Out Time")

```



## Question 2



**Question**:Do certain airlines lead to greater taxi out times?



**Answer**: Frontier and Southwest airlines have greater taxi-out time according to the below boxplot.

```{r}
ggplot(data = flights_df, mapping = aes(x = airline, 
                                 y = taxi_out, fill = airline)) +
  geom_boxplot() + 
  labs(title = "Boxplot of Taxi out time by Airlines",
       x = "Airline", 
       y = "Taxi Out Time")
```



## Question 3



**Question**:Are certain airlines prone to delays?



**Answer**: Based on average delay time, SkyWest airlines seem to be delayed much with average of 108.3 minutes compared to other airlines and the least delayed is Southwest airlines with average delay time of 49.7 minutes


```{r}
airport_delay <- delayedFlights_df %>% group_by(airline) %>% summarize(meanDep_delay = mean(dep_delay))

airport_delay 
```

## Question 4



**Question**:Are certain times of the day problematic?



**Answer**: 6-7 PM is the time with most number of flights operated which was problematic as there might have been waiting time for flights for take off or to land and use the run way.




```{r}

ggplot(data = delayedFlights_df, mapping = aes(x = dep_time, y = ..density..)) +
       geom_histogram(fill = "#006EA1", color = "white", bins = 15) + 
       labs(title = "Distribution of Flights in a day",
            x = "Departure Time",
            y = "Number of flights") + geom_density(fill = "black", alpha = 0.2)


```




## Question 5



**Question**:Which month has most number of delayed flights?



**Answer**: July 2016 has most number of delayed flights and June, December next to July which can be due to June, July falls in summer and December in winter. where all the months will have holidays for summer break and christmas.




```{r}
delayedFlightsCount <- delayedFlights_df %>% count(month, name = 'DelayedFlights',sort = TRUE)

delayedFlightsCount

ggplot(data = delayedFlightsCount, mapping = aes(x = month, y = DelayedFlights, fill= month)) +
       geom_bar(stat = "identity")  + 
       labs(title = "Delayed flights by month", x = "Month",
            y = "Number of delayed flights")
```

## Question 6



**Question**:How are airline, airport region and tax-in are related?



**Answer**:  American airlines in west region has highest average of 16.68 mins of tax-in and United airlines in middle atlantic region has lowest average of 3.95 mins of tax-in 




```{r}
flightDistance <- flights_df%>% group_by(airline, dest_airport_region) %>% summarize(meanTaxiIn = mean(taxi_in))

flightDistance
```


## Question 7



**Question**: How are arrival delay and distance affected airlines and destination region?



**Answer**: According to the below line graph, it is observed that if distance is more than 1500miles, arrival delay is around 60mins.


```{r}
airlinesDistance <- delayedFlights_df %>% group_by(airline, dest_airport_region) %>% summarize(AvgArrivalDelay = mean(arrival_delay), AvgDistance = mean(distance))
airlinesDistance


```


```{r}

ggplot(data = airlinesDistance, mapping = aes(x = AvgArrivalDelay, y = AvgDistance)) +
  geom_line(color = "#0072B2") +
  geom_point(color = "#0072B2") +
  theme_light() +  labs(title = "Boxplot of Arrival delay Vs distance",
       x = "Arrival delay", 
       y = "Distance")


```



## Question 8



**Question**: which weekday has more number of delayed flights?



**Answer**: Thursday has highest number of delayed flights with 893 , then comes Monday with 811 flights and the least count is on saturday with 578 flights




```{r}

WeekdayDemand <- delayedFlights_df%>% group_by(weekday) %>% summarize(delayedFlightsCount = n(), AvgTaxiIn = mean(taxi_in), AvgTaxiOut = mean(taxi_out) )

WeekdayDemand

```


# Summary of Results

Write an executive summary of your overall findings and recommendations to the executives at Dulles Airport. Think of this section as your closing remarks of a presentation, where you summarize your key findings and make recommendations on flight operations and strategies for minimizing flight delays.

Your executive summary must be written in a [professional tone](https://www.universalclass.com/articles/writing/business-writing/appropriate-tone-in-business-communications.htm), with minimal grammatical errors, and should include the following
sections:

1.  An introduction where you explain the business problem and goals of your data analysis

    -   What problem(s) is this company trying to solve? Why are they important
        to their future success?

    -   What was the goal of your analysis? What questions were you 
        trying to answer and why do they matter?


2.  Highlights and key findings from your Exploratory Data Analysis section

    -   What were the interesting findings from your analysis and
        **why are they important for the business**?

    -   This section is meant to **establish the need for your recommendations** 
        in the following section
        

3.  Your recommendations to the company

    -   Each recommendation must be supported by your data analysis results

    -   You must clearly explain **why** you are making each recommendation and which results 
        from your data analysis support this recommendation

    -   You must also describe the potential business impact of your recommendation:

        -   Why is this a good recommendation?

        -   What benefits will the business achieve?
      


## Executive Summary

Please write your executive summary below. If you prefer, you can type your summary in a text editor, such as Microsoft Word, and paste your final text here.

    --> To sustain operations, profitability, customer pleasure, and compliance with state and federal legislation, airports rely on accurate flight departure and arrival estimations. To maintain airport operations, flight performance, including departure and arrival delays, must be monitored, reported to the Federal Aviation Administration (FAA) on a regular basis, and minimized. If a flight is delayed by at least 15 minutes, the FAA considers it to be delayed.
    To improve their customer service and enhance business policies and strategies it is really important for the business to handle this way.
    
    --> My goal on this analysis is to improve the business by getting deeper into the data and find the important points like flight delays which plays a major role in service to work on for further improvement
    
    --> The questions that I was able to answer through my study are listed below, and these are the insights that assist firms solve difficulties and achieve business succes:
    
    Are flight delays affected by taxi-out time?
    Do certain airlines lead to greater taxi out times?
    Are certain airlines prone to delays?
    Are certain times of the day problematic?
    Which month has most number of delayed flights?
    How are airlines, airport region and tax-in are related?
    How are arrival delay and distance affected airlines and destination region?
    which weekday has more number of delayed flights?
    
    --> Findings for the questions are mentioned below:
    
    Flight delay doesn't seem related/ affected by taxi-out time because if looking from left to right, the taxi out time and airlines time is more around 250 on x-axis and 50 on y-axis which is not moving further which is observed from scatter plot.
    
    Frontier and Southwest airlines have greater taxi-out time according to the boxplot mentioned in question 2
    
    Based on average delay time, SkyWest airlines seem to be delayed much with average of 108.3 minutes compared to other airlines and the least delayed is Southwest airlines with average delay time of 49.7 minutes which was done through detailed summary.
    
    6-7 PM is the time with most number of flights operated which was problematic as there might have been waiting time for flights for take off or to land and use the run way which is observed from a histogram.
    
    July 2016 has most number of delayed flights and June, December next to July which can be due to June, July falls in summer and December in winter. where all the months will have holidays for summer break and christmas based on bar chart
    
    American airlines in west region has highest average of 16.68 mins of tax-in and United airlines in middle atlantic region has lowest average of 3.95 mins of tax-in which is observed by detailed summary.
    
    According to line graph, it is observed that if distance is more than 1500miles, arrival delay is around 60mins.
    
    Thursday has highest number of delayed flights with 893 , then comes Monday with 811 flights and the least count is on saturday with 578 flights based on detailed summary
    

  ---Recommendation of Dulles airport as per my analysis are listed below:
  
  The airport can reduce delay time if they make sure Frontier and southwest airlines to wait for less time at the gate and allow to the runway by scheduling the time of other flights accordingly.
  
  SkyWest airlines delay time is higher than other airlines which can be controlled by managing the flight distance, frequency of the flights and time of a particular day.
  
  As, there are more number of flights operated between 6-7 PM which causes flight delays, so changing the schedule of flight timing might reduce flight delay.
  
  As June, July and December months have vacation time every year, it makes busier than normal during these months. Increasing the frequency of flights in non peak hours might help the airport to operate in this peak months.
  
 
